### 2021-10-24 Simulation Deep Dive.Rmd

---
title: "Simulation Deep Dive"
author: "Paul Harmon"
date: "10/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("OutlierCompNew.R")
library(ggplot2)
library(readr)
library(tibble)
library(dplyr)
library(tidyr)
library(janitor)
library(magrittr)
library(knitr)
library(Rtsne)
library(robustHD)
library(GGally)
library(stringr)
library(plotly)
library(R.utils)
library(vegan)
library(mnormt)

correctInfluence = function(result_df){
  tmp <- tail(result_df,1)
  det_rate <- length(which(tmp$PValues <= 0.05))/nrow(tmp)
  return(det_rate)}

# Correctly Identified Non-Influential
#head(df3list[[1]], 60)

correctNonInfluence = function(result_df){
  tmp <- head(result_df,60)
  det_rate <- length(which(tmp$PValues > 0.05))/nrow(tmp)
  return(det_rate)}

spuriousInfluence = function(result_df){
  tmp <- result_df
  det_rate <- length(which(result_df$PValues <= 0.05))/nrow(tmp)
  return(det_rate)}

#then the non-influential 
meantest = function(df){
  mean(df$PValues)
}

```

# Intro

This document shows some of the results from a handful of different simulations. In general, we consider a situation where the data are arranged in 3 groups of varying width and variability. Each simulation fixes the sample size within groups and all groups are simulated from multivariate normals with 10 features. Additionally, only a SINGLE influential point is considered at present.

In most scenarios, 61 observations are used - except when no outliers are present.

Simulation scenarios:

+ Influential point "bounces around" within range of 3 groups
+ Influential point is "smaller" ~ 3 SD below average
+ Influential point is "larger" ~ 3 SD above average
+ No outliers (only 60 observations)

Pending Scenarios:

+ Wider range of influential points
+ Wider spread of groups
+ More dimensions than 10
+ Influence on only a subset of features

# Multiple Groups

## Data

```{r}
g1 <- rmnorm(n = 10, mean = rep(20, 20), diag(35, 20)) %>%
    t() %>%
    data.frame()
  g2 <- rmnorm(n = 10, mean = rep(40, 20), diag(35, 20)) %>%
    t() %>%
    data.frame()
  g3 <- rmnorm(n = 10, mean = rep(60, 20), diag(35, 20)) %>%
    t() %>%
    data.frame()

  yes_structure <- rbind(g1, g2, g3) %>%
    data.frame() %>%
    mutate(Group = c(rep("1", 20), rep("2", 20), rep("3", 20)))

  yes_structure_scale <- lapply(yes_structure[,1:10], scale) %>% as_tibble()

  ### adds outliers - but not as a single group. Here we have an outlier that includes a really big one, a really small observation, and a very abnormal one
  #add_outliers <- rbind(rnorm(10, 3, .5), rnorm(10, -3, .5), rnorm(10, c(rep(-2,3),rep(0,3),rep(2,3),0), .5)) %>% as_tibble()

  add_outliers <- rbind(rnorm(10, 0, 1)) %>% as_tibble()

  names(add_outliers) <- names(yes_structure_scale)

  yes_structure_final <- bind_rows(yes_structure_scale, add_outliers) 
  yes_structure_final$Group <- c(yes_structure$Group, rep("Outlier",1))

  ##### Visualize the data and store in a list
  datplot <- yes_structure_final %>% mutate(Id = 1:nrow(yes_structure_final), Alpha = ifelse(Group %in% 'Outlier', 1,0.8)) %>%  pivot_longer(1:10) %>% ggplot(aes(name, value, color = Group, group = Id, alpha = Alpha)) + geom_line() + geom_point() + ggtitle("Simulated Data with Outliers Added In") + guides(alpha = FALSE)
datplot
```

## Results

```{r}
mds1 <- readRDS("2021-10-16_simulation1_mds_list.RDS")
tsne1 <- readRDS("2021-10-16_simulation1_tsne_list.RDS")

#### Generate a table of values to assess efficacy of detections -----------------#
df2list <- tsne1
df3list  <- mds1
#note - these will give values per run - so we need to take a look across the simulations

#t-SNE
ci1 <- sapply(df2list, correctInfluence)
cni1 <- sapply(df2list, correctNonInfluence)

#MDS
ci2 <- sapply(df3list, correctInfluence)
cni2 <- sapply(df3list, correctNonInfluence)

#mean detection rates
mean(ci1)
mean(cni1)

mean(ci2)
mean(cni2)

```

# Larger

## Data

```{r}
 g1 <- rmnorm(n = 10, mean = rep(20, 20), diag(35, 20)) %>%
    t() %>%
    data.frame()
  g2 <- rmnorm(n = 10, mean = rep(40, 20), diag(35, 20)) %>%
    t() %>%
    data.frame()
  g3 <- rmnorm(n = 10, mean = rep(60, 20), diag(35, 20)) %>%
    t() %>%
    data.frame()

  yes_structure <- rbind(g1, g2, g3) %>%
    data.frame() %>%
    mutate(Group = c(rep("1", 20), rep("2", 20), rep("3", 20)))

  yes_structure_scale <- lapply(yes_structure[,1:10], scale) %>% as_tibble()

  ### adds outliers - but not as a single group. Here we have an outlier that includes a really big one, a really small observation, and a very abnormal one
  #add_outliers <- rbind(rnorm(10, 3, .5), rnorm(10, -3, .5), rnorm(10, c(rep(-2,3),rep(0,3),rep(2,3),0), .5)) %>% as_tibble()

  add_outliers <- rbind(rnorm(10, 3, .5)) %>% as_tibble()

  names(add_outliers) <- names(yes_structure_scale)

  yes_structure_final <- bind_rows(yes_structure_scale, add_outliers) 
  yes_structure_final$Group <- c(yes_structure$Group, rep("Outlier",1))

  ##### Visualize the data and store in a list
  datplot <- yes_structure_final %>% mutate(Id = 1:nrow(yes_structure_final), Alpha = ifelse(Group %in% 'Outlier', 1,0.8)) %>%  pivot_longer(1:10) %>% ggplot(aes(name, value, color = Group, group = Id, alpha = Alpha)) + geom_line() + geom_point() + ggtitle("Simulated Data with Outliers Added In") + guides(alpha = FALSE)
datplot
```

## Results

```{r}
mds1 <- readRDS("2021-10-17_simulationlarger_mds_list.RDS")
tsne1 <- readRDS("2021-10-17_simulationlarger_tsne_list.RDS")

#### Generate a table of values to assess efficacy of detections -----------------#
df2list <- tsne1
df3list  <- mds1
#note - these will give values per run - so we need to take a look across the simulations

#t-SNE
ci1 <- sapply(df2list, correctInfluence)
cni1 <- sapply(df2list, correctNonInfluence)

#MDS
ci2 <- sapply(df3list, correctInfluence)
cni2 <- sapply(df3list, correctNonInfluence)

#mean detection rates
mean(ci1)
mean(cni1)

mean(ci2)
mean(cni2)

```

# Smaller

## Data

```{r}
 g1 <- rmnorm(n = 10, mean = rep(20, 20), diag(35, 20)) %>%
    t() %>%
    data.frame()
  g2 <- rmnorm(n = 10, mean = rep(40, 20), diag(35, 20)) %>%
    t() %>%
    data.frame()
  g3 <- rmnorm(n = 10, mean = rep(60, 20), diag(35, 20)) %>%
    t() %>%
    data.frame()

  yes_structure <- rbind(g1, g2, g3) %>%
    data.frame() %>%
    mutate(Group = c(rep("1", 20), rep("2", 20), rep("3", 20)))

  yes_structure_scale <- lapply(yes_structure[,1:10], scale) %>% as_tibble()

  ### adds outliers - but not as a single group. Here we have an outlier that includes a really big one, a really small observation, and a very abnormal one
  #add_outliers <- rbind(rnorm(10, 3, .5), rnorm(10, -3, .5), rnorm(10, c(rep(-2,3),rep(0,3),rep(2,3),0), .5)) %>% as_tibble()

  add_outliers <- rbind(rnorm(10, -3, .5)) %>% as_tibble()

  names(add_outliers) <- names(yes_structure_scale)

  yes_structure_final <- bind_rows(yes_structure_scale, add_outliers) 
  yes_structure_final$Group <- c(yes_structure$Group, rep("Outlier",1))

  ##### Visualize the data and store in a list
  datplot <- yes_structure_final %>% mutate(Id = 1:nrow(yes_structure_final), Alpha = ifelse(Group %in% 'Outlier', 1,0.8)) %>%  pivot_longer(1:10) %>% ggplot(aes(name, value, color = Group, group = Id, alpha = Alpha)) + geom_line() + geom_point() + ggtitle("Simulated Data with Outliers Added In") + guides(alpha = FALSE)
datplot

```

## Results
```{r}
mds1 <- readRDS("2021-10-18_simulationsmaller_mds_list.RDS")
tsne1 <- readRDS("2021-10-18_simulationsmaller_tsne_list.RDS")

#### Generate a table of values to assess efficacy of detections -----------------#
df2list <- tsne1
df3list  <- mds1
#note - these will give values per run - so we need to take a look across the simulations

#t-SNE
ci1 <- sapply(df2list, correctInfluence)
cni1 <- sapply(df2list, correctNonInfluence)

#MDS
ci2 <- sapply(df3list, correctInfluence)
cni2 <- sapply(df3list, correctNonInfluence)

#mean detection rates
mean(ci1)
mean(cni1)

mean(ci2)
mean(cni2)

```

# No Outliers

## Data 
```{r}
 g1 <- rmnorm(n = 10, mean = rep(20, 20), diag(35, 20)) %>%
    t() %>%
    data.frame()
  g2 <- rmnorm(n = 10, mean = rep(40, 20), diag(35, 20)) %>%
    t() %>%
    data.frame()
  g3 <- rmnorm(n = 10, mean = rep(60, 20), diag(35, 20)) %>%
    t() %>%
    data.frame()

  yes_structure <- rbind(g1, g2, g3) %>%
    data.frame() %>%
    mutate(Group = c(rep("1", 20), rep("2", 20), rep("3", 20)))

  yes_structure_scale <- lapply(yes_structure[,1:10], scale) %>% as_tibble()

  ### No outliers - only 60 observations in this dataset

  yes_structure_final <- yes_structure_scale
  yes_structure_final$Group <- c(yes_structure$Group)

  ##### Visualize the data and store in a list
  datplot <- yes_structure_final %>% mutate(Id = 1:nrow(yes_structure_final), Alpha = ifelse(Group %in% 'Outlier', 1,0.8)) %>%  pivot_longer(1:10) %>% ggplot(aes(name, value, color = Group, group = Id, alpha = Alpha)) + geom_line() + geom_point() + ggtitle("Simulated Data with Outliers Added In") + guides(alpha = FALSE)
datplot
```

## Results
```{r}
#2021-10-18_simulationsmaller_mds_list

mds1 <- readRDS("2021-10-19_simulationNone_mds_list.RDS")
tsne1 <- readRDS("2021-10-19_simulationNone_tsne_list.RDS")

#### Generate a table of values to assess efficacy of detections -----------------#
df2list <- tsne1
df3list  <- mds1
#note - these will give values per run - so we need to take a look across the simulations

#t-SNE
ci1 <- sapply(df2list, spuriousInfluence)
cni1 <- 1-cni1

#MDS
ci2 <- sapply(df3list, spuriousInfluence)
cni2 <- 1 - cni2

#mean detection rates
mean(ci1)
mean(cni1)

mean(ci2)
mean(cni2)

```

# One Group

## Data

A look at the data, which is simulated from multivariate normal with a single group.

```{r}
 g1 <- rmnorm(n = 10, mean = rep(40, 60), diag(35, 60)) %>%
    t() %>%
    data.frame()

  yes_structure <- g1 %>%
    data.frame() %>%
    mutate(Group = c(rep("1", 60)))

  yes_structure_scale <- lapply(yes_structure[,1:10], scale) %>% as_tibble()

  ### adds outliers - but not as a single group. Here we have an outlier that includes a really big one, a really small observation, and a very abnormal one
  add_outliers <- rbind(rnorm(10, 3, .5), rnorm(10, -3, .5), rnorm(10, c(rep(-2,3),rep(0,3),rep(2,3),0), .5)) %>% as_tibble()
  names(add_outliers) <- names(yes_structure_scale)

  yes_structure_final <- bind_rows(yes_structure_scale, add_outliers) 
  yes_structure_final$Group <- c(yes_structure$Group, rep("Outlier",3))

  ##### Visualize the data and store in a list
  datplot <- yes_structure_final %>% mutate(Id = 1:nrow(yes_structure_final), Alpha = ifelse(Group %in% 'Outlier', 1,0.8)) %>%  pivot_longer(1:10) %>% ggplot(aes(name, value, color = Group, group = Id, alpha = Alpha)) + geom_line() + geom_point() + ggtitle("Simulated Data with Outliers Added In") + guides(alpha = FALSE)
datplot
```

## Results 
We see that the proportion of correctly-identified influential points seems to go down as we constrain the shape of the data to a single group.

```{r}
mds2 <- readRDS("2021-10-08_simulation2_mds_list.RDS")
tsne2 <- readRDS("2021-10-08_simulation2_tsne_list.RDS")

#### Generate a table of values to assess efficacy of detections -----------------#
df2list <- tsne2
df3list  <- mds2
#note - these will give values per run - so we need to take a look across the simulations

#t-SNE
ci1 <- sapply(df2list, correctInfluence)
cni1 <- sapply(df2list, correctNonInfluence)

#MDS
ci2 <- sapply(df3list, correctInfluence)
cni2 <- sapply(df3list, correctNonInfluence)

#mean detection rates
mean(ci1)
mean(cni1)

mean(ci2)
mean(cni2)

```

# Computation Time

```{r}
pt1 <- readRDS('perf_time_1017.RDS')
pt2 <- readRDS('perf_time_cmd_1017.RDS')

unlist(pt1)
#unlist(pt2)

pt1 %>% unlist() %>% as_tibble() %>% ggplot(aes(value)) + geom_histogram(bins = 10) + xlab("Minutes") + ggtitle("Computation Time - TSNE Runs") + theme_bw()

```